---
title: "Practical Machine Learning Project Write-Up"
output: html_document
---

Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 
Project Goal

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. 

```{r}
library(RCurl)
library(caret)
train_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
train_data <- read.csv(text=getURL(train_url), na.strings=c("", "NA", "<NA>"))
test_data <- read.csv(text=getURL(test_url), na.strings=c("", "NA", "<NA>"))
```

Then perform some exploratory data analysis
```{r, echo =FALSE, results='hide', message=FALSE, warning=FALSE}
library(Hmisc)
summary(train_data)
describe(train_data)
head(train_data)
colnames(train_data)
```
Then I started slicing the data. In the code below, I removed the first columns which included data and variables like name, timestamp, etc as they didn't appear to impact the classe or movements. Next I removed the columns that had NAs. Based on some exploratory data analysis, there appeared to be trend that the columns with NAs were mostly, if not all NAs. 
```{r}
trainDataSmall <- train_data[, -c (1:7)]
testDataSmall <- test_data[, -c (1:7)]
training_data <- trainDataSmall[ , !apply(is.na(trainDataSmall), 2, any)]
testing_data <- testDataSmall[ , !apply(is.na(testDataSmall), 2, any)]
```
Next, I partitioned the training set futher for cross-validation. I decided to keep 75% of the original training set to train the model on.
```{r}
inTrain <- createDataPartition(y=training_data$classe,
                               p=0.75, list=FALSE)
training <- training_data[inTrain,]
CVtesting <- training_data[-inTrain,]
```
Next I set the seed for replication. I then perform cross-validation while training the model on the new training set and predict outcomes on the data we set aside for cross-validation within the training set. The accuracy is 99.2% and an estimate of error rate of .67%
```{r}
library(randomForest)
library(caret)
set.seed(1000)
cvControl = trainControl(method='cv',number=5,repeats=2,returnResamp='none')
modelFit <- train(training$classe~.,data=training,method="rf", trControl=cvControl)
predictions <- predict(modelFit,newdata=CVtesting)

confusionMatrix(predictions,CVtesting$classe)
modelFit$finalModel
```
Then I run the model on the seperate testing data. When submitting the predicted answers, the predictions were all accurate.
```{r}
predictionsTest <- predict(modelFit,newdata=testing_data)
```
